Fission product decay chain equilibria in \zirconia

### 2019-01-25 manuscript.tex
Add Robin's corrections

This study follows previous work on defect equilibria in \zirconia\ to determine the oxide layer's effectiveness as a barrier to iodine \cite{kenichiodine2018}. It was found that the tetragonal phase of \zirconia\ is a greater barrier to iodine ingress than monoclinic \zirconia\ as the partial pressure of oxygen is increased. It is also known that tetragonal \zirconia\ will always be present on the inner surface of the cladding in significant quantities because it is self-stabilised by the stresses imposed as the oxide grows into the zirconium metal, in addition to compressive residual stresses induced by radiation damage. The iodine defect study, however, only informs us about one part of the SCC process. For a more holistic understanding, the life cycle of the iodine must be taken into account as well.     \\

@@

Nuclear fuel claddings have unique materials challenges associated with them owing to the highly active environment and creation of unstable isotopes. Corrosive species in the pin such as iodine can be produced directly as a result of fission of uranium fuel. While it is known that iodine plays a role in SCC, one must also consider that these iodine nuclei are unstable. Fission of uranium will produce iodine precursors, mainly unstable isotopes of tellurium. Both iodine and tellurium are relatively common fission products, with combined independent yields from thermal fission of U$_{235}$ above 5\% \cite{kennett1956mass, iodine129fissionyield, imanishi1976independent, iaeafissionyield, iodine132, amiel1975odd}. In addition, xenon and caesium are also common fission products, as shown in Figure \ref{fig:fissionyield}, with fission product masses firmly within the heavy section of the bi-modal fission product distribution curve.    \\

Nuclei produced during fission are typically neutron-rich, resulting in decay modes such as $\beta-$ or neutron emission. In the case of tellurium, the vast majority of unstable isotopes will decay into iodine, which then decays into xenon with varying half-lives depending on the isotope. The decay chain continues with xenon nuclei decaying into caesium, many isotopes of which have half lives measured in years. At this point, fuel is typically retired long before a significant quantity of caesium decays into barium. For this reason we only consider the elements tellurium through caesium in this study. It should also be noted that the majority of thermal fission events occur in the outer rim of the fuel pellet, and a fission product penetration depth of up to 8 $\mu$m in \zirconia\ \cite{degueldre2001behaviour} suggests a large degree of fission product implantation within the oxide. With each nuclear decay comes a change in the chemical and therefore physical behaviour of the atom with its immediate environment. For example, an iodine dopant in \zirconia\ may decay into xenon which will then have a significantly different thermodynamic equilibrium site from the one it inherited.   \\

Determining the effect of each of these elements in the oxide layer may provide information about the initiation of SCC in fuel cladding. We have therefore adopted a quantum-mechanical calculation approach to model the behaviour of the decay chain elements tellurium through caesium within tetragonal phase zirconia. \\

>>>>

Nuclear fuel claddings have unique materials challenges owing to the highly active environment in which they operate and the creation of unstable isotopes. Corrosive species in the pin such as iodine, produced directly as a result of uranium fission are known to play a role in SCC [REF]. However, iodine nuclei produced from fission are unstable. Fission of uranium produces iodine precursors, mainly unstable isotopes of tellurium. Both iodine and tellurium are relatively common fission products, with combined independent yields from thermal fission of U$_{235}$ above 5\% \cite{kennett1956mass, iodine129fissionyield, imanishi1976independent, iaeafissionyield, iodine132, amiel1975odd}. In addition, xenon and caesium are also common fission products, as shown in Figure \ref{fig:fissionyield}, with fission product masses firmly within the heavy section of the bi-modal fission product distribution curve.    \\

Previous work on defect equilibria in \zirconia\ considered the internal oxide layer's effectiveness as a barrier to iodine \cite{kenichiodine2018}. It was found that the tetragonal phase of \zirconia\ is a better barrier to iodine ingress than monoclinic \zirconia , especially at higher oxygen partial pressures. It is also established that tetragonal \zirconia\ will always be present on the inner surface of the cladding in significant quantities because it is self-stabilised by the stresses imposed as the oxide grows into the zirconium metal, in addition to compressive residual stresses induced by radiation damage. The iodine defect study, however, only informs us about one part of the SCC process. For a more holistic understanding, the life cycle of the iodine must also be taken into account.     \\

@@

Nuclei produced during fission are typically neutron-rich, resulting in decay modes such as $\beta-$ or neutron emission. In the case of tellurium nuclei, most decay into iodine, which then decays into xenon with different half-lives depending on the isotope. The decay chain continues with xenon nuclei decaying into caesium, many isotopes of which have half lives measured in years. At this point, conventional power reactor fuel is retired long before a significant proportion of the caesium decays into barium. For this reason we only consider the elements tellurium through caesium in this study. \\

The majority of thermal fission events occur in the outer rim of a fuel pellet. Given a fission product penetration depth of up to 8 $\mu$m in \zirconia\ \cite{degueldre2001behaviour}, a large proportion of fission products implant within the oxide. \\

With each nuclear decay comes a change in the chemical and therefore physical behaviour of the atom with its immediate environment. For example, an iodine dopant in \zirconia\ may decay into xenon, which may then have a significantly different thermodynamic equilibrium site preference from the one it inherited. Determining the behaviour of each of these elements in the oxide layer may provide information about the initiation of SCC in fuel cladding. We have therefore adopted a quantum-mechanical calculation approach to model the behaviour of the decay chain elements tellurium through caesium within tetragonal phase zirconia. \\
